CUNY SPSS Data Science Math Homework 1

### 2018-07-18 data-science-math/Homework 1.Rmd
Additional work

$$f(x) = \frac{-8}{x^2}$$
$$f'(x) = \frac{16}{x^3}$$
Verify with R:
```{r}
f5 <-function(x){-8/(x**2)}
Deriv(f5)
```

@@

$$g'(x) \frac{5}{3x^\frac{2}{3}}=\frac{5}{3}\cdot\frac{1}{x^\frac{2}{3}}$$
```{r}
f6 <- function(x){5*x**(1/3)}
Deriv(f6)

@@

$$y'=\frac{-9x^\frac{1}{8}}{4}=\frac{-9}{4}\cdot x^\frac{1}{8}$$
```{r}
f7 <- function(x){-2*x**(9/8)}
Deriv(f7)

@@

$$f'(x) = \frac{4}{x^3}(-5x+9) + -5(2x^{-2} + 1) = \frac{36}{x^3} - \frac{20}{x^2} + \frac{10}{x^2} - 5=\frac{36}{x^3}-\frac{10}{x^2}-5$$
Checking my math with R:
```{r}
f10 <- function(x){(-2*x**-2+1)*(-5*x+9)}
Deriv(f10)
# Evaluating at 10
f10_prime <- function(x){(36/x**3)-(10/x**2)-5}
f10_prime(10)
f10_prime <- Deriv(f10)
f10_prime(10)
```

@@

Simplified to 
$$f'(x) = -\frac{42x^\frac{5}{2} + 25x^3 - 5}{2\sqrt{x}\cdot (x^3 + 1)^2}$$
Again checking with R:
```{r}
f11 <- function(x){(5*x**(1/2)+7)/(x**3+1)}
Deriv(f11)
# Evaluate at 5
fprime11 <- function(x){-1*(42*x**(5/2)+25*x**3-5)/((2*x**(1/2))*((x**3+1)**2))}
fprime11(5)
fprime11 <- Deriv(f11)
fprime11(5)
```

@@

Multiplying by 550 and simplifying
$$f'(x) = \frac{1100t\sqrt{t^2+15} -\frac{550t^3}{\sqrt{t^2+15}}}{t^2+15}=\frac{550t(t^2+30)}{(t^2+15)^\frac{3}{2}}$$
```{r}
f13 <- function(t){(550*t**2)/((t**2+15)**(1/2))}
Deriv(f13)
f13_prime <- function(t){(550*t*(t**2+30))/((t**2+15)**(3/2))}
f13_prime(3)
f13_prime <- Deriv(f13)
f13_prime(3)

```

@@

Given $N(t) = 1000(6+0.1t)^\frac{1}{2}$
$$N(3) = 1000(6+0.1(3))^\frac{1}{2}=39690$$

>>>>

$$f(x) = \frac{-8}{x^2}$$
$$f'(x) = \frac{16}{x^3}$$

Verify with R:
```{r}
f5 <-function(x){-8/(x**2)}
Deriv(f5)
```

@@

$$g'(x) \frac{5}{3x^\frac{2}{3}}=\frac{5}{3}\cdot\frac{1}{x^\frac{2}{3}}$$

Verify with R:
```{r}
f6 <- function(x){5*x**(1/3)}
Deriv(f6)

@@

$$y'=\frac{-9x^\frac{1}{8}}{4}=\frac{-9}{4}\cdot x^\frac{1}{8}$$

Verify with R:
```{r}
f7 <- function(x){-2*x**(9/8)}
Deriv(f7)

@@

$$f'(x) = \frac{4}{x^3}(-5x+9) + -5(2x^{-2} + 1) = \frac{36}{x^3} - \frac{20}{x^2} + \frac{10}{x^2} - 5=\frac{36}{x^3}-\frac{10}{x^2}-5$$

Verify with R:
```{r}
f10 <- function(x){(-2*x**-2+1)*(-5*x+9)}
Deriv(f10)
# Evaluating at 10
f10_prime <- function(x){(36/x**3)-(10/x**2)-5}
f10_prime(10)
f10_prime <- Deriv(f10)
f10_prime(10)
```

@@

Simplified to 
$$f'(x) = -\frac{42x^\frac{5}{2} + 25x^3 - 5}{2\sqrt{x}\cdot (x^3 + 1)^2}$$

Again checking with R:
```{r}
f11 <- function(x){(5*x**(1/2)+7)/(x**3+1)}
Deriv(f11)
# Evaluate at 5
fprime11 <- function(x){-1*(42*x**(5/2)+25*x**3-5)/((2*x**(1/2))*((x**3+1)**2))}
fprime11(5)
fprime11 <- Deriv(f11)
fprime11(5)
```

@@

Multiplying by 550 and simplifying
$$f'(x) = \frac{1100t\sqrt{t^2+15} -\frac{550t^3}{\sqrt{t^2+15}}}{t^2+15}=\frac{550t(t^2+30)}{(t^2+15)^\frac{3}{2}}$$

Verifying with R:
```{r}
f13 <- function(t){(550*t**2)/((t**2+15)**(1/2))}
Deriv(f13)
f13_prime <- function(t){(550*t*(t**2+30))/((t**2+15)**(3/2))}
f13_prime(3)
f13_prime <- Deriv(f13)
f13_prime(3)
```

@@

$$N(t) = 1000(6+0.1t)^\frac{1}{2}=1000\sqrt{0.1t+6}$$

#### Step 1
$$N(3) = 1000(6+0.1(3))^\frac{1}{2}=39690$$

#### Step 2
Let $u=0.1t+6$ so $\sqrt{u}$ becomes $\frac{1}{2\sqrt{u}}$ because of the power rule

and due to the chain rule $u'=0.1$.  Substitute

$$1000\cdot N'(t) = \frac{0.1\cdot0.5}{\sqrt{0.1t+6}}=\frac{0.05}{\sqrt{0.1t+6}}$$
$$ N'(t)=\frac{1000\cdot 0.05}{\sqrt{0.1t+6}}=\frac{50}{\sqrt{0.1t+6}}$$
$$N'(5)=\frac{50}{\sqrt{0.1(5)+6}}=\frac{50}{\sqrt{6.5}}=19.6116135138\approx20	$$
20 is roughly how many additional people attend in the 5th week of the season

Checking my math with R:
```{r}
f14 <- function(t){1000*(6+0.1*t)**(1/2)}
Deriv(f14)
f14_prime <- function(t){50/(6.5**(1/2))}
f14_prime(5)
f14_prime <- Deriv(f14)
f14_prime(5)
```

### Question 15

$$3x^3+4y^3=77$$

$$\frac{d}{dx}(3x^3+4y^3)=\frac{d}{dx}(77)$$

$$\frac{d}{dx}(3x^3)+\frac{d}{dx}(4y^3)=0$$
$$9x^2+12y^2\frac{dy}{dx}=0$$

$$12y^2\frac{dy}{dx}=-9x^2$$

$$\frac{dy}{dx}=\frac{-9x^2}{12y^2} = -\frac{3x^2}{4y^2}$$

The slope of the tangent line at (3,-1) is

$$-\frac{3(3)^2}{4(-1)^2}=-\frac{3(9)}{4(1)}=-\frac{7}{4}$$

### Question 16

$$f(x)=\frac{x+3}{x-8}$$

Differentiate using the quotient rule where $f(x)=x+3$ and $g(x)=x-8$

$$f'(x)=\frac{(1)(x-8)-(1)(x+3)}{(x-8)^2}=\frac{(x-8)-(x+3)}{(x-8)^2}=\frac{x-8-x-3}{(x-8)^2}=-\frac{11}{(x-8)^2}$$

Checking my work with R:
```{r}
f16 <- function(x){(x+3)/(x-8)}
Deriv(f16)
f16_prime <- function(x){-11/((x-8)^2)}
f16_prime(20)
f16_prime <- Deriv(f16)
f16_prime(20)
```

$$f''(x)= \frac{22}{(x-8)^3}$$

The critical point is when x=8.  f(x) is decreasing when x < 8 and increasing when x > 8.

### Question 17

$$F(t) = 14 + \frac{367t^2}{t^2+100}$$

$$367 \cdot F'(t) = \frac{t^2}{(t^2+100)^2}=\frac{(2t(t^2+100))(-2t^2\cdot t)}{(t^2+100)^2}=\frac{2t(t^2+100)(-2t^3)}{(t^2+100)^2}$$

Multiply it by 367

$$f'(x)=367 \cdot \frac{2t(t^2+100)(-2t^3)}{(t^2+100)^2}=\frac{734t(t^2+100)-734t^3}{(t^2+100)^2}=\frac{73400t}{(t^2+100)^2}$$

In the interest of turning my assignment on time some of the steps have been excluded

$$f''(x)=-\frac{73400(3t^2-100)}{(t^2+100)^3}$$

$F'(t) > 0$ from $(0,\infty)$ and $F'(t) < 0$ from $(-\infty, 0)$ which doesn't make sense in this case.

$$\lim_{t\to\infty} F(t) = 381^\circ F$$

### Question 18

$$f(x)=\frac{x^3}{3340000}-\frac{7x^2}{9475}+\frac{42417727x}{1265860000}+\frac{1}{33}$$

Again in the interest of time I will skip the steps of the calculation:

$$f'(x) = \frac{3x^2}{3340000}-\frac{14x}{9475}+\frac{42417727}{1265860000}$$

$$0 = \frac{3x^2}{3340000}-\frac{14x}{9475}+\frac{42417727}{1265860000}$$

```{r}
f18 <- function(x){}
f18_prime <- function(x){(3*x**2/3340000)-(14*x/9475)+(42417727/1265860000)}
x <- seq(0, 30, by=.01)
y <- f18_prime(x)
i <- stats::uniroot(f18_prime,c(-30,30))
plot(y~x,type="l", col="blue")
abline(h=0, col="black")
abline(v=0, col="black")
abline(v=i$root, col="red")
points(i$root, 0, col="red", pch=21)
```

```{r}
i$root
```

### Question 19

$$f(x) = 7x^2+28x-35$$

$$f'(x) = 14x+28$$

$$0 = 14x+28 $$

$$-28 = 14x$$

$$x=-\frac{28}{14}=-2$$

$$f(-2) = (7(-2)^2)+(28(-2))-35 = 28-56-35=-63$$

$$\mbox minimum = (-2,-63)$$

Checking with R:
```{r}
f19 <- function(x){7*x**2+28*x-35}
f19_prime <- Deriv(f19)
i <- stats::uniroot(f19_prime,c(-30,30))
i$root
f19(i$root)
```

```{r}
x <- seq(-30, 30, by=.01)
y <- f19(x)
i <- stats::uniroot(f19_prime,c(-30,30))
plot(y~x,type="l", col="blue")
abline(h=0, col="black")
abline(v=0, col="black")
points(i$root, 0, col="red", pch=21)
```

### Question 20

$$f(x)=-6x^3+27x^2+180x$$

$$f'(x)=-18x^2+54x+180$$

$$0=-18x^2+54x+180$$

$$0=-18(x^2-3x-10)$$

$$x^2-3x-10=0$$

$$(x+2)(x-5)=0$$

$$x=-2 and 5$$
